---
title: "Tidy Tuesday Screencasts - Squirrels in NYC"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Squirrels in NYC
- [Video](https://www.youtube.com/watch?v=6GV9sAD6Pi0), [Stopped here](https://youtu.be/6GV9sAD6Pi0?t=2377)
- [Github repo](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-10-29)


# What have we done here?
* breaking down Central park into hectares
* use by_hectare data to visualizae number of SQ
* use `summarize_at` function to get percentages across many metrics per hectare 
* create figure on NYC shapefile
* create Shiny app to make many plots across cutoffs and variables


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
theme_set(theme_light())
nyc_squirrels <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-29/nyc_squirrels.csv")
```

```{r}
nyc_squirrels %>% count(zip_codes, sort = T)
```

Plot Lat and Long to deterime if this is central park. Also, each point is a squirrel
```{r}
nyc_squirrels %>%
  ggplot(aes(long, lat)) + geom_point()
```

Can we simplify the plot since it looks too busy? Let's summarize data on the hectare level. Since hectare is a string, let's average the Lat and Long of all points in a Hectare and keep track of number of Squirrels in each Hectare.
```{r}
by_hectare = nyc_squirrels %>%
  filter(!is.na(primary_fur_color)) %>%   # this is for next block of code
  group_by(hectare) %>%
  summarize(long=mean(long),
            lat = mean(lat),
            n = n(),
            pct_gray = mean(primary_fur_color == "Gray", na.rm = T))  # for next block of code

by_hectare %>% ggplot(aes(long, lat, size = n)) + geom_point() +
  theme_void()
```

Hypothesis: fur color is dependent on the location. There is a slight reduction in Gray squirrels in smaller latitudes.
```{r}
nyc_squirrels %>% count(primary_fur_color, sort = T)

# Only show hectares with at least 10 S
by_hectare %>% filter(n >= 10) %>%
  ggplot(aes(long, lat, size = n, color = pct_gray)) +
  geom_point()

# Seems like there is a trend
by_hectare %>% filter(n >= 10) %>%
  ggplot(aes(lat, pct_gray)) +
  geom_point() + 
  geom_smooth()

# Test it with a logistic regression
by_hectare %>% mutate(n_gray = round(pct_gray * n)) %>%
  glm(cbind(n_gray, n - n_gray) ~ lat, data = ., family = "binomial") %>%  # <-- I don't understand how/why he chose these as the y-variables. I thought you need Y to be a binary factor (see: https://www.statmethods.net/advstats/glm.html)
  summary()
```
S may be more likely to be gray at higher north in the park you go


Exploring fur colors
```{r}
nyc_squirrels %>%
  count(primary_fur_color, highlight_fur_color, sort =T)

# how many highlight fur colors do we have
nyc_squirrels %>%
  count( highlight_fur_color, sort =T)
```


```{r}
nyc_squirrels %>%
  count(approaches, indifferent, runs_from, sort = T)
```

Does a S run away based on latitude?
```{r}
glm(runs_from ~ lat, data = nyc_squirrels, family="binomial") %>%     
  summary()

by_hectare = nyc_squirrels %>%
  add_count(hectare) %>%
  # above_ground mutate was added at ~57min
  mutate(above_ground = !is.na(nyc_squirrels$location) & nyc_squirrels$location =='Above Ground') %>%
  group_by(hectare, n) %>%  # shows how many S per hectare
  summarize_at(vars(long, lat, kuks:runs_from, indifferent, ends_with("ing"), above_ground), mean) %>% # summarizing multivars with same function
  ungroup()

# without any scale_XX_XX coloring
by_hectare %>%
  filter(n>=10) %>%
  ggplot(aes(long, lat, size = n, color = runs_from)) +
  geom_point() +
  theme_void() +
  labs(color = "% of S run", size = "# of S",
       title = "S in NW of Central Park are more likely to run away")

# change color scaling of % of S Run (similar to Red Blue highlight in Excel). Add a 'mid' color because default is white
# see link for scales::percent
#    https://rdrr.io/cran/scales/man/label_percent.html
by_hectare %>%
  filter(n>=10) %>%
  ggplot(aes(long, lat, size = n, color = runs_from)) +
  geom_point() +
  theme_void() +
  scale_color_gradient2(low = "blue", high = "red", mid="pink",
          labels=scales::percent, midpoint = .3) + # creates a 3 color divergent gradient
  labs(color = "% of S", size = "# of S",
       title = "S in NW of Central Park are more likely to run away")
```

```{r}
library(ggmap)
#?ggmap

#get_map(location = c(lon=-73.98, lat=40.77))

#by_hectare %>%
#  ggmap() +
#  geom_point(aes(long, lat, color = runs_from))
```



```{r}
library(sf)

central_park_sf = read_sf("CentralAndProspectParks")

by_hectare %>%
  filter(n>=10) %>%
  ggplot() +
  geom_sf(data = central_park_sf) +
  geom_point(aes(long,lat, size=n,color=runs_from)) +
  theme_void() +
  scale_color_gradient2(low = "blue", high = "red", mid="pink",
          labels=scales::percent, midpoint = .3) + # creates a 3 color divergent gradient
  labs(color = "% of S", size = "# of S",
       title = "S in NW of Central Park are more likely to run away") +
  coord_sf(datum = NA)
```


```{r}
nyc_squirrels %>% count(other_activities, sort=T)
nyc_squirrels %>% select(ends_with("ing"))

# Gather column names ending in "ing" into a column called 'activity' then label that 'activity' as true or false in column 'doing'. The Gather command will use the values in the '*ing' columns for the 'doing' column
nyc_squirrels %>% gather(activity, doing, ends_with("ing")) %>% select(activity, doing)


nyc_squirrels %>% gather(activity, doing, ends_with("ing")) %>% filter(doing == T) %>% count(activity, sort=T)

```


```{r}
# Look at histograms of variables
# Need to 'gather' these varaibles to go from Wide to Long for ggplot
by_hectare %>% gather(variable, value, kuks:foraging) %>%
  filter(n>=10) %>%
  ggplot(aes(value)) + geom_histogram() + facet_wrap(~ variable)
```


```{r}
nyc_squirrels %>% count(location)  # from this he added a Mutate to by_hectare to add % of SQ that are 'Above Ground'
```


```{r}
sessionInfo()
```

