---
title: "Ten Tremendous Tricks"
output: html_document
---

# Ten Tidyverse Tricks using Diamond dataset
[Video](https://www.youtube.com/watch?v=NDHSBUN_rVU). Thank you David Robinson!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadlibrary, warning=FALSE}
library(tidyverse)
library(broom) # for 'tidy'
library(purrr) # for 'map'
```

### 1. count()
Display a sorted frequency table of number diamonds in each Clarity group

```{r count}
diamonds %>% count(clarity, sort=T, name = "Number")  # sortedfreq
diamonds %>% count(clarity, wt=price, sort=T)  # sum of Price for each Clarity group
```

### 2. creating variables in count()
Show number of diamonds in each $1000 price range

```{r createvarscount}
diamonds %>% count(thous = 1000 * price %/% 1000) %>% ggplot(aes(thous, n)) +
  geom_line()
```

### 3. add_count()
Add column to diamonds showing the number of diamonds in each Clarity group

```{r addcount}
diamonds %>% add_count(clarity)
# select Clarity with > 3000 diamonds
diamonds %>% add_count(clarity) %>% filter(n>3000)
```

### 4. summarize() w/ a list column
Test the NULL hypo that price is 4000 regardless of clarity  t.test mu=4000 [video](https://youtu.be/NDHSBUN_rVU?t=576)

```{r summarizeListColumn}
ttests = diamonds %>% group_by(clarity) %>% summarize(t_test = list(t.test(price, mu=4000)))
ttests %>% mutate(tidied = map(t_test, tidy)) %>% unnest(tidied) %>%
  ggplot(aes(estimate, clarity)) + geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high))
```

### 5. fct_reorder + geom_col + coord_flip

```{r}
# Reorder Clarity groups descending by total number of diamonds (on the fly!)
diamonds %>% count(clarity) %>% mutate(clarity = fct_reorder(clarity, n)) %>% 
  ggplot(aes(clarity, n)) + geom_col() + coord_flip()
```

### 6. fct_lump()
Before using 'lump'. This shows all Clarity groups. But lets say you want to only show the top 3 frequenct Clarity groups and LUMP all others into an Other category (see next plot) 
```{r}
diamonds %>% ggplot (aes(clarity, price)) + geom_boxplot() + coord_flip()
```

After using 'fct_lump'. This lumps Clarity groups into Other if they are not the top 3 most frequent Clarity groups. Then reorder Clarity factor levels descending by 'price' using 'fct_reorder'
```{r}
diamonds %>% mutate(clarity=fct_lump(clarity,3), clarity=fct_reorder(clarity, price)) %>% ggplot (aes(clarity, price)) + geom_boxplot() + coord_flip()
```



```{r sessioninfo}
sessionInfo()
```

