---
title: "Ten Tremendous Tricks"
output: html_document
---

# Ten Tidyverse Tricks using Diamond dataset
[Video](https://www.youtube.com/watch?v=NDHSBUN_rVU). Thank you David Robinson!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadlibrary, warning=FALSE}
library(tidyverse)
library(broom) # for 'tidy'
library(purrr) # for 'map'
```

### 1. count()
Display a sorted frequency table of number diamonds in each Clarity group

```{r count}
diamonds %>% count(clarity, sort=T, name = "Number")  # sortedfreq
diamonds %>% count(clarity, wt=price, sort=T)  # sum of Price for each Clarity group
```

### 2. creating variables in count()
Show number of diamonds in each $1000 price range

```{r createvarscount}
diamonds %>% count(thous = 1000 * price %/% 1000) %>% ggplot(aes(thous, n)) +
  geom_line()
```

### 3. add_count()
Add column to diamonds showing the number of diamonds in each Clarity group

```{r addcount}
diamonds %>% add_count(clarity)
# select Clarity with > 3000 diamonds
diamonds %>% add_count(clarity) %>% filter(n>3000)
```

### 4. summarize() w/ a list column
Test the NULL hypo that price is 4000 regardless of clarity  t.test mu=4000 [video](https://youtu.be/NDHSBUN_rVU?t=576)

```{r summarizeListColumn}
ttests = diamonds %>% group_by(clarity) %>% summarize(t_test = list(t.test(price, mu=4000)))
ttests %>% mutate(tidied = map(t_test, tidy)) %>% unnest(tidied) %>%
  ggplot(aes(estimate, clarity)) + geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high))
```

### 5. fct_reorder + geom_col + coord_flip

```{r fct_reorder}
# Reorder Clarity groups descending by total number of diamonds (on the fly!)
diamonds %>% count(clarity) %>% mutate(clarity = fct_reorder(clarity, n)) %>% 
  ggplot(aes(clarity, n)) + geom_col() + coord_flip()
```

### 6. fct_lump()
Before using 'lump'. This shows all Clarity groups. But lets say you want to only show the top 3 frequenct Clarity groups and LUMP all others into an Other category (see next plot) 
```{r fct_lump}
diamonds %>% ggplot (aes(clarity, price)) + geom_boxplot() + coord_flip()
```

After using 'fct_lump'. This lumps Clarity groups into Other if they are not the top 3 most frequent Clarity groups. Then reorder Clarity factor levels descending by 'price' using 'fct_reorder'
```{r fct_lump2}
diamonds %>% mutate(clarity=fct_lump(clarity,3), clarity=fct_reorder(clarity, price)) %>% ggplot (aes(clarity, price)) + geom_boxplot() + coord_flip()
```

### 7. Log scales scale_x/y_log10()
Alot of data is Log normal so let's check the Carat distribution before and after Logging

```{r logscales_before}
diamonds %>% ggplot(aes(carat)) + geom_histogram()# + scale_x_log10()
```

```{r logscales_after}
diamonds %>% ggplot(aes(carat)) + geom_histogram() + scale_x_log10()
```

Perhaps(??) you get a better prediction of Price with Log Carat
```{r logscales_prediction}
diamonds %>% ggplot(aes(carat, price)) + geom_point() + scale_x_log10()
```

### 8. Crossing()
[video](https://www.youtube.com/watch?v=NDHSBUN_rVU)
From tidyr package
- returns a tibble
- no conversion of strings to factors
- can cross two tibbles together
```{r crossing}
x = crossing(a = 1:10, b= c("a","b"), c=c("w","x","y","z"))
```

```{r crossing_mathcurves}
# total genes
# total genes DNArep
# total sample genes (i.e. Diff Regulated Genes)
# total sample genes DNArep
total_genes = 10000                 # analogous to total_dogs
total_genesDNArep = c(100,200,300)  # total genes DNArep varies as research progresses
                                    # analogous to name_total
total_sample = seq(200,1000,25)     # samples with different number of DEG genes
                                    # analogous to breed_total
crossing(total_genesDNArep = total_genesDNArep, total_sample=total_sample) %>% 
  mutate (maxpvalue = 1 - phyper(0, m = total_genesDNArep, n = total_genes - total_genesDNArep, k=total_sample)) %>%
  ggplot(aes(total_sample, maxpvalue, color=factor(total_genesDNArep))) + geom_line()

```
```{r crossing_augment}
#TBD
```

```{r crossing_tidysimulation}
#TBD
```

### 9. separate()
Opposite of tidyr 'unite'
```{r}
mtcars
x = mtcars %>% tbl_df() %>% unite(mpg, cyl, col = "temp") %>% mutate(mpgcyl = temp) %>% select(temp, mpgcyl)
x
x %>% separate(mpgcyl, c("mpg", "cyl"), sep = "_")
```

### 10. extract()
From tidyr. Useful for pulling out patterns embedded in a field. Here I'm splitting 'cut' into the first and second word.
```{r}
diamonds
diamonds %>% tidyr::extract(cut, c("firstword","secondword"), "(.*) (.*)", convert=T)
```



```{r sessioninfo}
sessionInfo()
```

